---
title: "Coregulation_across_Tissues_Species"
author: "Natasha Santhanam"
date: "2/14/2022"
output: html_document
---

````{r}
library(tidyverse)
library(devtools)
library(broom)
library(data.table)
library(RSQLite)
library(Hmisc)
"%&%" = function(a,b) paste(a,b,sep="")
dir <- "/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/"
set.seed(777)
```

1. Correlation between genes within each tissue and then
calculate correlation of the correlation between genes across tissues

```{r}
filelist <- list.files("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan", pattern = "predict.txt", full.names = TRUE)
orth.rats <- read_tsv(dir %&% "expression/ortholog_genes_rats_humans.tsv", col_names = TRUE)
```

Generate Correlation Matrices for 100 rats - compare predicted Expression across tissues
```{r}
names <- read_tsv(filelist[1]) %>% select(c(FID))

for(i in 1:100) {
  id = sample(names$FID, size = 1)
  tempo <- data.frame(gene = as.character())
  for(fila in filelist) {
    name <- substr(fila, 58,59)
    tis <- fread(fila) %>% filter(FID == id) %>% pivot_longer(!c(FID, IID), names_to = "gene", values_to = name) %>% select(-c(FID, IID))
    tempo <- full_join(tempo, tis,  by = "gene")
  } 
  tempo <- tempo %>% mutate(var = apply(tempo[,-1], 1, var)) %>% na.omit()
  saveRDS(tempo, dir %&% "prediXcan/GREx_comp/" %&% id %&% ".GREx.mat.RDS")
}
```


Check heatmap of some individuals
```{r}
i1 <- readRDS("/Users/natashasanthanam/Downloads/00077E6712.cor.mat.RDS")
i2 <- readRDS("/Users/natashasanthanam/Downloads/00077E7788.cor.mat.RDS")

melted_i1 <- melt(i1, na.rm = TRUE)
melted_i2 <- melt(i2, na.rm = TRUE)

p1= ggplot(data = melted_i1, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),) + ggtitle("00077E6712")
p2= ggplot(data = melted_i2, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),) + ggtitle("00077E7788")

ggarrange(p1, p2, ncol=2)
```

2.  Co-regulation Shared Across Species

```{r}
gene_ids <- data.frame(id = fread(filelist[1]) %>% select(-c(FID, IID)) %>%  colnames())
for(fila in filelist[2:length(filelist)]) {
  df <- data.frame(id = fread(fila) %>% select(-c(FID, IID)) %>%  colnames())
  gene_ids <- inner_join(gene_ids, df, by = "id")
}
```


```{r}
gtf <- fread(dir %&% "Box_files/gtf.txt", header = TRUE)
gtf <-  gtf[match(tempo$id, gtf$Gene),]
 
for(i in 1:20) {
  tempo <- gtf %>% filter(Chr == i) %>% select(c(Gene))
  if(nrow(tempo) == 0 ) {
  i = i+1  
  }
  else {
  df <- data.frame(row = as.character(), column = as.character())
  for(fila in filelist) {
    tis <- substr(fila, 58,59)
    expr <- as.data.frame(fread(fila) %>% select(-c(FID, IID)))
    expr <- expr[,intersect(tempo$Gene, colnames(expr))]
    res2<-rcorr(as.matrix(expr[,]))
  d2 <- flattenCorrMatrix(res2$r, res2$P)
    colnames(d2)[3] = tis
    colnames(d2)[4] = paste("p", tis, sep = "_")
    df <- full_join(df, d2, by = c("row", "column"))
  }
  saveRDS(df, dir %&% "prediXcan/GREx_comp/cor_genes_per_chr/chr" %&% i %&% ".RDS" )
  }
}
```


Save correlation of coregulation across tissues
```{r}
coreg.dir <- "/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan/GREx_comp/cor_genes_per_chr"
filelist <- list.files(coreg.dir, pattern = ".RDS", full.names = TRUE)

for(fila in filelist) {
  tempo <- readRDS(fila) %>% select(c(row, column, Ac, Il, Pl, Lh, Vo)) %>%
  i <- substr(fila, 89, str_length(fila)- 4)
  cor.mat <-  cor(tempo[,3:7])
  saveRDS(cor.mat, dir %&% "prediXcan/GREx_comp/cor_genes_per_chr/cor_coreg_chr" %&% i %&% ".RDS")
}
```


Graph of Correlation of Coregulation across tissues

```{r}
data.dir <- "/Users/natashasanthanam/CRI/"
filelist <- list.files(data.dir, pattern="cor_coreg", full.names = TRUE)
corr_coreg <- list()

for(fila in filelist) {
  i <- match(fila, filelist)
  corr_coreg[[i]] <- readRDS(fila)
}
```

Generate Coregulation in Humans (GTEx)

First generate predicted expression in Framingham using GTEx models
```{bash}
conda activate imlabtools
export METAXCAN=/gpfs/data/im-lab/nas40t2/natasha/GTEX_Analysis/MetaXcan/software
export GENO=/gpfs/data/im-lab/nas40t2/Data/dbGaP/Transcriptome/Framingham/imputed_hrc1.1
export MODEL=/gpfs/data/im-lab/nas40t2/Data/PredictDB/GTEx_v8/models_v1/eqtl/ctimp
export RESULTS=/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/PTRS_weights/PGP
export DATA=/gpfs/data/im-lab/nas40t2/natasha/GTEX_Analysis/predixcan

printf "Predict expression\n\n"

python3 $METAXCAN/Predict.py \
--model_db_path $MODEL/ctimp_$TISSUE.db \
--model_db_snp_key varID \
--vcf_genotypes $GENO/chr*.dose.vcf.gz  \
--vcf_mode genotyped \
--liftover $DATA/hg19ToHg38.over.chain.gz  \
--on_the_fly_mapping METADATA "chr{}_{}_{}_{}_b38" \
--prediction_output $RESULTS/Framingham__$TISSUE.predict.txt \
--prediction_summary_output $RESULTS/Framingham__$TISSUE.summary.txt \
--verbosity 9 \
--throw

TISSUE=Brain_Cerebellum
```


```{r flatten corr mat function, eval=FALSE}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```


Calculate Coregulation between genes in GTEx
```{r}
gtf <- fread("/gpfs/data/im-lab/nas40t2/natasha/GTEX_Analysis/annotations_gencode_v26.tsv", header = TRUE)

 for(i in 1:20) {
  tempo <- gtf %>% filter(chromosome == paste("chr", i, sep="")) %>% select(c(gene_id))
  expr <- as.data.frame( fread("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan/GREx_comp/Framingham_less_mem__Brain_Cerebellum.predict.txt")) %>% select(-c(FID, IID)) 
 colnames(expr) = sapply(strsplit(colnames(expr), "\\."), `[`, 1)
    expr <- expr[, intersect(tempo$gene_id, colnames(expr))]
   res2<-rcorr(as.matrix(expr[,]))
  d2 <- flattenCorrMatrix(res2$r, res2$P)
saveRDS(d2, dir %&% "prediXcan/GREx_comp/cor_GTEx_genes_per_chr/GTEx_chr" %&% i %&% ".RDS" )
 }

```

```{r}
filelist <- list.files("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan/GREx_comp/cor_GTEx_genes_per_chr/", pattern = ".RDS", full.names = TRUE)
human_genes <- data.frame()

for(fila in filelist) {
  tempo <- readRDS(fila)
  human_genes <- rbind(tempo, human_genes)
}

summary(human_genes$GTEx)
#Mean is 0.00266

filelist <- list.files("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan/GREx_comp/cor_genes_per_chr", "chr", full.names = TRUE)[1:20]
rat_genes <- data.frame()
for(fila in filelist) {
  tempo <- readRDS(fila)
  rat_genes <- rbind(tempo, rat_genes)
}
```


Check if Coregualtion is preserved across species
```{r}
orth.rats <- read_tsv("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/expression/ortholog_genes_rats_humans.tsv")
all_genes <- data.frame()
for(i in 1:20) {
  rat.mat <- readRDS(dir %&% "prediXcan/GREx_comp/cor_genes_per_chr/chr" %&% i %&% ".RDS")
  human.mat <- readRDS(dir %&% "prediXcan/GREx_comp/cor_GTEx_genes_per_chr/GTEx_chr" %&% i %&% ".RDS")
  
  rat.mat <- rat.mat %>% filter(row %in% orth.rats$rnorvegicus_homolog_ensembl_gene & column %in% orth.rats$rnorvegicus_homolog_ensembl_gene) 
  rat.mat$row = orth.rats[match(rat.mat$row, orth.rats$rnorvegicus_homolog_ensembl_gene),1]$ensembl_gene_id
  rat.mat$column = orth.rats[match(rat.mat$column, orth.rats$rnorvegicus_homolog_ensembl_gene),1]$ensembl_gene_id
  
  human.mat <- human.mat %>% filter(row %in% rat.mat$row | row %in% rat.mat$column)
  if(nrow(human.mat) != 0) {
    rat.mat <- rat.mat %>% filter(row %in% human.mat$row | row %in% human.mat$column)
    human.mat <- full_join(human.mat, rat.mat, by = c("row", "column"))
    all_genes <- rbind(human.mat, all_genes)

  }
}
all_genes <- readRDS("/Users/natashasanthanam/Downloads/coregu_across_species_larger.RDS")
```


Plot Coregulation across species
```{r}
all_genes <- all_genes %>% dplyr::select(-c(row, column))
pairs(all_genes)
cor.test(all_genes$GTEx, all_genes$Ac)

ggplot(all_genes, aes(Vo, GTEx)) + geom_point() + ylim(0, 1)
ggplot(all_genes, aes(Vo, Ac)) + geom_point() + ylim(0, 1)
```
